---
title: "Water and Life Expectancy (Unicef)"
author: "Emilie GU"
format: html
execute:
  echo: false
  warning: false
  message: false
editor: visual
---

# **Water, the source of life ?**

## **Accessibility to clean drinking water sources could improve life expectancy in developing countries**

-   Developing countries in the South have less proportion of the population using clean water sources compared to developed countries in the North.

```{r Map}

library(tidyverse)
library(ggplot2)
library(maps)
library(janitor)

unicef_indicator_1final <- read_csv2("~/Desktop/unicef_indicator_1final.csv") %>%
  clean_names()
unicef_metadata_final <- read_csv2("~/Desktop/unicef_metadata_final.csv")

#map visualisation
Improved_water_2020 <- unicef_indicator_1final %>% 
  filter(year == 2020)
names(unicef_indicator_1final)[7] ="obs_value"

map_world <- map_data("world")
map_water_2020 <- full_join(map_world, Improved_water_2020, by = c("region" = "country"))

ggplot(data = map_water_2020) +
  aes(x = long, y = lat, group = group, fill = obs_value_percent) +
  geom_polygon() +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "lightgrey") +
  labs(
    title = "Proportion of the population using improved drinking water source in 2020",
    subtitle = "Countries in grey have no data collected",
    x = "Longitude",
    y = "Latitude",
    fill = "obs value (%) in 2020"
  ) +
  theme_classic()+
  theme(plot.title = element_text(size=13, face = "bold"))
```

-   Usage of clean water sources can explain **63%** of changes in life expectancy. More precisely, the higher the proportion of the population using of improved drinking water sources, the longer the life expectancy for a country's population.

```{r Scatter plot}
#scatter plot with regression line
Life_expectancy <- full_join(unicef_indicator_1final, unicef_metadata_final, by = c("year" = "year", "country"="country"))
Life_expectancy <- Life_expectancy[,-c(2,3,4,8,9,10,11,12,13,14,15,16,17,19)]
Life_expectancy <- na.omit(Life_expectancy)

x <- Life_expectancy$obs_value
y <- Life_expectancy$`Life expectancy at birth, total (years)`
plot(Life_expectancy$obs_value, Life_expectancy$`Life expectancy at birth, total (years)`, 
     main = 'Regression life expectation and proportion of population using
     improved drinking water sources', cex.main=1,
     xlab = 'Obs value (%)', ylab = 'Life expectancy at birth (years)', col="lightblue",cex=0.5)
abline (lm (y~ x, data = Life_expectancy), col= "darkblue") 

```

-   An improvement of 7% on average since 2000 but still not enough !

```{r Time series}

#Time-series chart
MeanObsvalue<-unicef_indicator_1final%>%
  group_by(year) %>%
  summarise(m_obsvalue = mean(obs_value))

ggplot(data = MeanObsvalue) +
  aes(x = year, y = m_obsvalue, color = m_obsvalue) +
  geom_line() +
  labs(
    x = "Year (from 2000 to 2020)",
    y = "Average Obs value in the world",
    title = "The average proportion of population using improved drinking water sources 
over the years."
  ) +
  theme_light()+
  theme(plot.title = element_text(size=13, face = "bold", hjust = 0.5))
```

-   The average life expectancy is also lower in developing countries in comparison to developed countries.

```{r Bar chart, echo = FALSE}
#bar chart
MeanLifeExp <- Life_expectancy %>% 
  group_by(country) %>% 
  summarise(m_lifeExp = mean(`Life expectancy at birth, total (years)`))
  MeanLifeExp <-arrange(MeanLifeExp, m_lifeExp)
  MeanLifeExp <-MeanLifeExp [c(1, 124), ]
  
ggplot (data = MeanLifeExp) +
  aes(x = country, y = m_lifeExp, fill = country) +
  geom_bar(stat="identity")+
  scale_fill_manual(values=c("lightblue",
                             "darkblue"))+
labs(
    x = "Country : lowest vs highest",
    y = "Average life expectancy",
    title = "Country with the lowest vs highest life expectancy at birth.",
    color="lightblue"
  ) +
  geom_hline(yintercept=72.3)+
  theme_light()+
  theme(plot.title = element_text(size=13, face = "bold", hjust = 0.5))
```
